Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-31 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Last Revision: Last modification (by Jeff) on the paper ... now it is up to Kostas
updates: adding figures for the qualitative validation -- Section 4.2 ... text to come ...

%%%%%%%%%%%% 
\subsection{Initial Model Validation}\label{section:results_initial_model_validation}
Numerical simulations \citep[based on lab experiments due to][]{evans_1994,dawe_2008} were conducted to demonstrate the model's capability to capture viscous crossflow during immiscible displacement in heterogeneous porous media. The 2D domain, shown in Fig.~\ref{fem_cv_represent_a}, is $4$ $\times$ $2$ unit-length and fully saturated with fluid 2 (VR=1). Fluid 1 is injected from the left-hand side of the domain with constant velocity of $u_{x}=1$. Boundary conditions also include no-flux across upper and lower borders. The domain consists of four regions in which each quarter is represented by a permeability value, Fig.~\ref{fem_cv_represent_a}(a).

A two-phase immiscible flow along homogeneous and parallel layers of contrasting petrophysical properties (\ie permeability) is initially simulated. During fluid displacement, crossflow between adjacent layers often occur due to viscous, capillary and/or gravitational forces in discontinuous spatial permeability distribution. Such preferential flows are shown in Figs.~\ref{fem_cv_represent_a}(b-c). In this case, crossflow behaviour from first to fourth quadrants of the domain is in good qualitative agreement with experiments conducted by \citet{dawe_2008} (see Figs. 5-6 of this reference).

>>>>

%%%%%%%%%%%% 
\subsection{Initial Model Validation}\label{section:results_initial_model_validation}
Numerical simulations \citep[based on lab experiments due to][]{evans_1994,dawe_2008} were conducted to demonstrate the model's capability to capture viscous crossflow during immiscible displacement in heterogeneous porous media. The 2D domain, shown in Fig.~\ref{DaweGrattoniExperiment_a}(a), is $4$ $\times$ $2$ unit-length and fully saturated with fluid 2 (VR=1). Fluid 1 is injected from the left-hand side of the domain with constant velocity of $u_{x}=$8.33$\times$10$^{-3}$. Boundary conditions also include no-flux across upper and lower borders. The domain consists of four regions in which each quarter is represented by a permeability value.

A two-phase immiscible flow along homogeneous and parallel layers of contrasting permeability is initially simulated. During fluid displacement, crossflow between adjacent layers often occur due to viscous, capillary and/or gravitational forces in discontinuous spatial permeability distribution. Such preferential flows are shown in Figs.~\ref{DaweGrattoniExperiment_b}. In this case, crossflow behaviour from first to fourth quadrants of the domain is in good qualitative agreement with experiments conducted by \citet{dawe_2008}, Fig.~\ref{DaweGrattoniExperiment_a}.
